---
title: "Functions"
author: "Jill Reiner"
date: "12/2/2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyverse)
library(tidyr)
library(stringr)
library(janitor)
```

let's write some functions

get defensive players:
```{r}
get_defense <- function(full_tracking) {
  full_tracking %>%
    filter(is_defender == 1, team != "football") %>%
    select(game_id, play_id, frame_id, event, nfl_id, display_name, position, defensive_team,
           yards_to_go, yardline_number, yardline_side, possession_team, epa, pass_success,
           x, y, s, a, dis, o, dir, x_rel_to_ball, y_rel_to_ball, dist_to_ball, x_rel_to_qb, y_rel_to_qb, dist_to_qb, x_rel_to_targ, y_rel_to_targ, dist_to_targ, x_targ, y_targ) %>%
    rename(nfl_id_def = nfl_id,
           name_def = display_name,
           pos_def = position,
           x_def = x,
           y_def = y,
           s_def = s,
           a_def = a,
           dis_def = dis,
           o_def = o,
           dir_def = dir,
           x_rel_to_ball_def = x_rel_to_ball,
           y_rel_to_ball_def = y_rel_to_ball, 
           dist_to_ball_def = dist_to_ball,
           x_rel_to_qb_def = x_rel_to_qb,
           y_rel_to_qb_def = y_rel_to_qb,
           dist_to_qb_def = dist_to_qb,
           x_rel_to_targ_def = x_rel_to_targ,
           y_rel_to_targ_def = y_rel_to_targ,
           dist_to_targ_def = dist_to_targ)
}
```

get offensive players:
```{r}
get_offense <- function(full_tracking) {
  
  full_tracking %>%
    filter(is_defender == 0, team != "football") %>%
    select(game_id, play_id, frame_id, event, nfl_id, display_name, position, is_targ_receiver, route,
           x, y, s, a, dis, o, dir, x_rel_to_ball, y_rel_to_ball, dist_to_ball, x_rel_to_qb, y_rel_to_qb, dist_to_qb) %>%
    rename(nfl_id_off = nfl_id,
           name_off = display_name,
           pos_off = position,
           x_off = x,
           y_off = y,
           s_off = s,
           a_off = a,
           dis_off = dis,
           o_off = o,
           dir_off = dir,
           x_rel_to_ball_off = x_rel_to_ball,
           y_rel_to_ball_off = y_rel_to_ball, 
           dist_to_ball_off = dist_to_ball,
           x_rel_to_qb_off = x_rel_to_qb,
           y_rel_to_qb_off = y_rel_to_qb,
           dist_to_qb_off = dist_to_qb)
}
```

distance from offensive players to each defensive player:
```{r}
get_dist_from_o_to_d <- function(full_tracking){
  
  offense <- get_offense(full_tracking)
  defense <- get_defense(full_tracking)
  
  dist_to_closest_opp <- offense %>%
    left_join(defense, by = c("game_id",
                              "play_id",
                              "frame_id",
                              "event")) %>%
    mutate(x_from_closest_opp = x_off - x_def,
           y_from_closest_opp = y_off - y_def,
           dist_from_opp = sqrt((x_from_closest_opp)^2 + (y_from_closest_opp)^2)) %>%
    group_by(name_off, play_id, frame_id) %>%
    mutate(closest_opp_distance = min(dist_from_opp)) %>%
    ungroup()
}
```

```{r}
get_dist_from_o_to_d_sec <- function(full_tracking){
  
  offense <- get_offense(full_tracking)
  defense <- get_defense(full_tracking)
  
  dist_to_closest_opp <- offense %>%
    left_join(defense, by = c("game_id",
                              "play_id",
                              "frame_id",
                              "event")) %>%
    mutate(x_from_closest_opp = x_off - x_def,
           y_from_closest_opp = y_off - y_def,
           dist_from_opp = sqrt((x_from_closest_opp)^2 + (y_from_closest_opp)^2)) %>%
    group_by(name_off, play_id, frame_id) %>%
    mutate(sec_closest_opp_distance = min(dist_from_opp[dist_from_opp!=min(dist_from_opp)])) %>%
    ungroup()
}
```

distance from offensive players to each defensive back:
```{r}
get_dist_from_o_to_backs <- function(full_tracking){
  
  offense <- get_offense(full_tracking)
  defense <- get_defense(full_tracking)
  
  backs <- defense %>%
    filter(pos_def %in% c("CB", "DB", "SS", "FS", "S"))
  
  dist_to_closest_opp <- offense %>%
    left_join(backs, by = c("game_id",
                              "play_id",
                              "frame_id",
                              "event")) %>%
    mutate(x_from_closest_opp = x_off - x_def,
           y_from_closest_opp = y_off - y_def,
           dist_from_opp = sqrt((x_from_closest_opp)^2 + (y_from_closest_opp)^2)) %>%
    group_by(name_off, play_id, frame_id) %>%
    mutate(closest_opp_distance = min(dist_from_opp)) %>%
    ungroup()
}
```

```{r}
min_opp_dists <- get_dist_from_o_to_d(full_tracking) %>%
  group_by(game_id, play_id, frame_id, nfl_id_off) %>%
  slice(which.min(dist_from_opp)) %>%
  select(-dist_from_opp) %>%
  ungroup()
```

```{r}
sec_min_opp_dists <- get_dist_from_o_to_d_sec(full_tracking) %>%
  group_by(game_id, play_id, frame_id, nfl_id_off) %>%
  slice(which.min(sec_closest_opp_distance)) %>%
  ungroup()
```

```{r}
second_closest_opp <- sec_min_opp_dists %>%
  select(game_id, play_id, frame_id, event, nfl_id_off, sec_closest_opp_distance)

min_opp_dists <- min_opp_dists %>%
  left_join(second_closest_opp, by = c("game_id", "play_id", "frame_id", "event", "nfl_id_off"))
```

```{r}
min_back_dists <- get_dist_from_o_to_backs(full_tracking) %>%
  group_by(game_id, play_id, frame_id, nfl_id_off) %>%
  slice(which.min(dist_from_opp)) %>%
  select(-dist_from_opp) %>%
  ungroup()
```

good
```{r}
min_back_dists_snap <- min_back_dists %>%
  mutate(rn = row_number()) %>%
  filter(event == "ball_snap",
         pos_off %in% c("WR"))
```

good
```{r}
min_opp_dists_snap <- min_opp_dists %>%
  mutate(rn = row_number()) %>%
  filter(event == "ball_snap",
         pos_off %in% c("RB", "FB", "HB", "TE"))
```

ok so for RB type players and TEs we're just using the closest opp at any position,

for WRs we don't want linebackers to be flagged as closest opp at snap so it is closest back or safety
```{r}
min_dists_snap <- min_back_dists_snap %>%
  bind_rows(min_opp_dists_snap) %>%
  arrange(rn)
```

same thing for pf?
```{r}
min_back_dists_pf <- min_back_dists %>%
  mutate(rn = row_number()) %>%
  filter(event == "pass_forward",
         pos_off %in% c("WR"))
```

```{r}
min_opp_dists_pf <- min_opp_dists %>%
  mutate(rn = row_number()) %>%
  filter(event == "pass_forward",
         pos_off %in% c("RB", "FB", "HB", "TE", "WR"))
```

```{r}
min_dists_pf <- min_back_dists_pf %>%
  bind_rows(min_opp_dists_pf) %>%
  arrange(rn)
```

```{r}
min_dists <- min_dists_snap %>%
  bind_rows(min_opp_dists_pf) %>% #min_dists_pf
  arrange(rn)
```


